¥documentclass[11pt,letterpaper]{article}
¥usepackage[margin=1in]{geometry}
¥usepackage{amsmath,amssymb,amsthm}
¥usepackage{hyperref}
¥usepackage{cleveref}
¥usepackage{booktabs}

¥numberwithin{equation}{section}
¥renewcommand{¥theequation}{S14.¥arabic{equation}}

¥title{Supplement S14: Quark Charges, Color, and Mass Ratios ¥¥
¥large Supporting Manuscript ¥S17}
¥author{UDT Collaboration}
¥date{}

¥begin{document}
¥maketitle

¥begin{abstract}
We derive the quark electric charges, the number of colors, and the
inter-generation mass ratios from the angular structure on $S^2$.
The color multiplicity $N_c = (2¥ell+1) = 3$ is the orbital degeneracy.
Quark charges follow from rank-2 parity blocks: $Q_d = -1/3$,
$Q_u = +2/3$.  Four inter-generation mass ratios are each $2¥pi^2$
times a sector factor, all sub-$2¥%$.  Verified by
¥texttt{analysis/20¥_quark¥_masses.py}.
¥end{abstract}

¥tableofcontents

%----------------------------------------------------------------------
¥section{Color as orbital orientation}
%----------------------------------------------------------------------

The Dirac equation on $S^2$ decomposes into angular channels labeled
by $¥kappa$.  At each $|¥kappa|$, the orbital angular momentum
$¥ell = 1$ provides a 3-dimensional representation space with basis
states $|m_¥ell¥rangle$ for $m_¥ell = -1, 0, +1$.  These three
orientations are identified with the three color states:
¥begin{equation}
  N_c = 2¥ell + 1 = 3¥,.
¥end{equation}
The $¥mathfrak{su}(3)$ algebra that governs color interactions closes
exactly on this $¥ell = 1$ subspace (Supplement‾S16).

%----------------------------------------------------------------------
¥section{Quark electric charges}
%----------------------------------------------------------------------

The rank-2 tensor operators on the $¥ell = 1$ space have a parity
block structure.  The sign counting within each rank-2 block gives:
¥begin{align}
  Q_d &= -¥frac{1}{N_c} = -¥frac{1}{3}¥,,¥¥
  Q_u &= +¥frac{2}{N_c} = +¥frac{2}{3}¥,.
¥end{align}
The total $R$-ratio:
¥begin{equation}
  R(u,d) = N_c(Q_u^2 + Q_d^2) = 3 ¥times ¥frac{5}{9} = ¥frac{5}{3}¥,,
¥end{equation}
matches the Standard Model exactly.

The charge assignment $Q_u = +2/3$ also follows from the bridge
identity: $¥sqrt{B_1/B_3} = ¥sqrt{(1/18)/(1/8)} = ¥sqrt{4/9} = 2/3$.

%----------------------------------------------------------------------
¥section{Inter-generation mass ratios}
%----------------------------------------------------------------------

The four ratios between adjacent generations follow a pattern:
each is $2¥pi^2$ times a sector factor from the $j$--$¥pi$--$¥ell$
bridge transition.

¥begin{center}
¥begin{tabular}{lclrc}
¥toprule
Ratio & Formula & Value & Expt. & Error ¥¥
¥midrule
$m_s/m_d$ & $2¥pi^2$ & $19.74$ & $20.0$ & $-1.3¥%$ ¥¥
$m_c/m_u$ & $6¥pi^4$ & $584.0$ & $587.5$ & $-0.6¥%$ ¥¥
$m_b/m_s$ & $9¥pi^2/2$ & $44.41$ & $44.8$ & $-0.8¥%$ ¥¥
$m_t/m_c$ & $14¥pi^2$ & $138.2$ & $136.0$ & $+1.6¥%$ ¥¥
¥bottomrule
¥end{tabular}
¥end{center}

The common factor $2¥pi^2$ is the baseline inter-generation step.
The sector factors $(1, 3¥pi^2, 9/4, 7)$ each involve the
multiplicities $(2, 3, 5, 7)$:
¥begin{itemize}
  ¥item $m_s/m_d$: baseline, no sector factor.
  ¥item $m_c/m_u = 6¥pi^4 = (2¥ell+1)(2j+1) ¥times (2¥pi^2)^2$:
    spin-orbital multiplied by squared step.
  ¥item $m_b/m_s = (9/4)¥pi^2 = [(2¥ell+1)/(2j+1)]^2 ¥times ¥pi^2$:
    squared orbital/spin ratio.
  ¥item $m_t/m_c = 14¥pi^2 = 2(2|¥kappa_¥mathrm{max}|+1) ¥times ¥pi^2$:
    twice the orbit size.
¥end{itemize}

¥subsection{Base quark masses}

The lightest quarks have masses set by the selection rule for rank-2
bilinears --- squared multiplicity:
¥begin{align}
  m_u &= (2j+1)^2¥,m_e = 4¥,m_e = 2.04‾¥text{MeV}
    ¥qquad(-5.4¥%)¥,,¥¥
  m_d &= (2¥ell+1)^2¥,m_e = 9¥,m_e = 4.60‾¥text{MeV}
    ¥qquad(-1.5¥%)¥,.
¥end{align}
These are approximate because $¥overline{¥text{MS}}$ scheme masses at
$¥mu = 2$‾GeV carry renormalization scheme dependence at the $¥sim5¥%$
level for the lightest quarks.

%----------------------------------------------------------------------
¥section{Verification}
%----------------------------------------------------------------------

All results verified by ¥texttt{analysis/20¥_quark¥_masses.py}.
Output: ¥texttt{data/generated/20¥_quark¥_masses.json}.

¥end{document}
